---
title: "02.1.2_adding_fish_var.Rmd"
author: "Jagadeesh Puvvula"
date: "2023-11-01"
output: pdf_document
---

```{r}
library(pacman)
p_load(tidyverse, readxl, janitor)
```

# regroup into 1=no consumed; 2: 1-3 times a month; 3: greater than three times a month
#HOME legend
#1= Almost every day (daily)
#2= 4 to 6 times a week (16-24 times a month)
#3= 1 to 3 times a week (4-12 times a month)
#4= 1 to 3 times a month
#5= less than once a month
#6= not at all

#MIREC legend
#0 = "None" - h6
#1 = ">0 and <=1/month" - h5
#2 = ">1 and <=2/month" - h6
#3 = ">2 and <=3/month" - h6
#4 = ">3 and <=4/month"
#5 = ">4 and <=5/month"
#6 = ">5 and <=6/month"
#7 = ">6 and <=7/month"
#8 = ">7 and <=8/month"
#9 = ">8 and <=9/month"
#10 = ">9/month"
```{r}
home_fish<- read_csv("E:/BBK17/pj/data_2023apr/home_fish_var.csv") |>
  mutate(fish_int_pm = case_when(
    fish_consumption_cd == 6 ~ 1,
    fish_consumption_cd %in% c(4, 5) ~ 2,
    fish_consumption_cd %in% c(1, 2, 3) ~ 3,
    TRUE ~ NA_integer_),
    subject_id = as.character(subject_id)) |>
  select(subject_id, fish_int_pm)

################
mirec_fish<- read_xlsx("E:/BBK17/8. BBK17_extra MIREC DVs_Oct26 2023/BBK17_MIREC_MOTHER_DVs_Yes_to_MIREC_BBK_Sept28 2023.xlsx") |>
  clean_names() |>
  select(subject_id, v1_general_fish_mthly_c) |>
  mutate(fish_int_pm = case_when(
    v1_general_fish_mthly_c == 0 ~ 1,
    v1_general_fish_mthly_c %in% c(1, 2, 3) ~ 2,
    v1_general_fish_mthly_c %in% c(4, 5, 6, 7, 8, 9, 10) ~ 3,
    TRUE ~ NA_integer_
  ))|>
  select(subject_id, fish_int_pm)
```

#append two datasets and export
```{r}
fish_intake<- bind_rows(home_fish, mirec_fish) |>
  mutate(fish_int_pm = replace_na(fish_int_pm, 2))
```

#add fish to final dataset
```{r}
dt<- read_csv("E:/BBK17/pj/data_2023apr/obs_w_outcome/wppsi_f_analy.csv")|>
  mutate(pb = if_else(cohort == "1", pb * 10, pb))


dt_w_fish<- left_join(dt, fish_intake, by="subject_id")
```

#export final dataset (As suggested by Aimin, renamed variables on Nov 27, 2023)
```{r}
dt_w_fish<- read_csv("E:/BBK17/pj/data_2023apr/obs_w_outcome/wppsi_f_analy.csv") |>
  rename( "bcep" = "bcetp",
         "bdcipp" = "bdcpp",
         "dnbp" = "dbup")

write_csv(dt_w_fish, "E:/BBK17/pj/data_2023apr/obs_w_outcome/wppsi_f_analy.csv")
```




