---
title: "03.5.1_bkmr-all"
author: "Jagadeesh Puvvula"
date: "2023-02-20"
output: pdf_document
---

```{r, echo=FALSE, message=FALSE}

library(pacman)
pacman::p_load(tidyverse, janitor, bkmr, Matrix, fields, bindrcpp)
path <- "E:/BBK17/pj/data_2023apr/obs_w_outcome/"

dat<- read_csv(paste0(path, "wppsi_f_analy.csv"))

```

```{r, echo=FALSE, message=FALSE}
dat<- dat |>
  mutate_if(is.numeric, round, 4) |>
  clean_names() |>
  rename(Pb=pb, Hg=hg, DMA=dma, DDE= dde, PBDE_47 = pbde47,
         PCB_118=pcb118, PCB_138=pcb138, PCB_153=pcb153,
         PCB_180=pcb180, PFHxS=pfhxs, PFOA=pfoa, PFOS=pfos,
         BCEP=bcep, BDCIPP=bdcipp, DNBP=dnbp, DPhP=dphp,
         TCS=tcs, BPA=bpa, MBP=mbp, MBZP=mbzp, MCPP=mcpp,
         sigma_DEHP=dehp, MEP=mep, MIBP=mibp, di_Ethyl_OP=op_de,
         di_Methyl_OP=op_dm, B_PB=b_pb, M_PB=m_pb, P_PB=p_pb, city=center,
         mom_edu_cat=edu3)|>
  mutate(race_bin = ifelse(race == "white", 1, 2))|>
  select(c(Pb, Hg, DMA, DDE, PBDE_47, PCB_118, PCB_138, PCB_153, PCB_180,
          PFHxS, PFOA, PFOS, BCEP, BDCIPP, DNBP, DPhP, TCS, BPA, MBP, MBZP,
          MCPP, sigma_DEHP, MEP, MIBP, di_Ethyl_OP, di_Methyl_OP, 
          B_PB, M_PB, P_PB, cotinine, home_score_total, mom_age, 
          cohort, city, sex, race_bin,  mom_edu_cat, parity_n, fish_int_pm,
           wppsi_fsiq, wppsi_viq, wppsi_piq,)) 
```

#for 29 chemicals
```{r}
chemicals_29<- c("Pb","Hg", "DMA",
              "DDE", "PBDE_47","PCB_118","PCB_138", "PCB_153","PCB_180",
              "PFHxS", "PFOA", "PFOS",
              "BCEtP", "BDCPP", "DBuP", "DPhP",
              "TCS", "BPA",
              "di_Ethyl_OP", "di_Methyl_OP",
              "MBP","MBZP","MCPP", "sigma_DEHP","MEP","MIBP",
              "B_PB", "M_PB", "P_PB")
chem_group_29<- c(rep(1,times=3), rep(2,times=6), rep(3, times=3),
               rep(4,times=4), rep(5,times=2), rep(6,times=2),
               rep(7,times=6), rep(8,times=3))
outcomes<-c("wppsi_fsiq","wppsi_viq", "wppsi_piq")
covariates<- c("cotinine", "home_score_total", "mom_age", "cohort",
                           "city", "sex", "race_bin", "mom_edu_cat", "parity_n", "fish_int_pm")

# for all chemicals select 1:32
dat <- dat |>
  select(1:42) |>
  drop_na()|>
  mutate_at(vars(1:30), ~log10(.+0.0000001)) |>
  drop_na() |>
  mutate_all(as.numeric)

dat_mirec <- dat |> filter(cohort=="2")
```

#grpreg LASSO with neg coef for at least one outcome: 15 chem
```{r}
chemicals_15<- c("Pb", "PFHxS", "P_PB", "di_Ethyl_OP",
                 "BDCPP", "DPhP", 
                 "TCS", "BPA",
                 "PBDE_47","PCB_138","PCB_153",
                 "MBP","MBZP","MCPP", "MEP")

chem_group_15<- c(rep(1,times=4), rep(2,times=2), rep(3, times=2),
               rep(4,times=3),rep(5,times=4))
outcomes<-c("wppsi_fsiq","wppsi_viq", "wppsi_piq")
covariates<- c("cotinine", "home_score_total", "mom_age", "cohort",
                           "city", "sex", "race_bin", "mom_edu_cat", "parity_n" )

# for all chemicals select 1:32
dat <- dat |>
  select(c(chemicals_15, covariates, outcomes)) |>
  drop_na()|>
  mutate_at(vars(1:17), ~log10(.+0.0000001)) |>
  drop_na() |>
  mutate_all(as.numeric)

table(dat$sex, dat$cohort)


dat_mirec <- dat |> filter(cohort=="2")
```


#grpreg LASSO with pos coef for at least one outcome: 14 chem
```{r}
chemicals_14<- c("Hg","DMA",
                 "DDE","PCB_118","PCB_180",
                 "PFOA","PFOS",
                 "BCEtP","DBuP",
                 "sigma_DEHP","MIBP",
                 "di_Methyl_OP",
                 "B_PB","M_PB")

chem_group_14<- c(rep(1,times=2), rep(2,times=3), rep(3, times=2),
               rep(4,times=2),rep(5,times=2),rep(6,times=1),rep(7,times=2))
outcomes<-c("wppsi_fsiq","wppsi_viq", "wppsi_piq")
covariates<- c("cotinine", "home_score_total", "mom_age", "cohort",
                           "city", "sex", "race_bin", "mom_edu_cat", "parity_n" )

# for all chemicals select 1:32
dat <- dat |>
  select(c(chemicals_14, covariates, outcomes)) |>
  drop_na()|>
  mutate_at(vars(1:15), ~log10(.+0.0000001)) |>
  drop_na() |>
  mutate_all(as.numeric)

table(dat$sex, dat$cohort)


dat_mirec <- dat |> filter(cohort=="2")
```

-------------------------------------------

=================================================================================
#BKMR loop
#Non-linearlity modeled using #knots = (15/sample size)*100
#convergence tested at 100 iterations
# TUNE nd value in knots100 object and iter in kmbayes function

#Run model (select either sex or cohort to TRUE: Convergence issue)
#for all 29 chemicals
```{r}
bkmr_sex(data=dat, 
          folder_path = "E:/BBK17/pj/data_2023apr/results/bkmr/29_chem", 
          ln_mixture_vars= chemicals_29,
          groups=chem_group_29,
          outcome_vars= outcomes, 
          covariate_vars= covariates, 
          include_sex = TRUE)
#gender sepcific estimates only for mirec
bkmr_sex(data=dat_mirec, 
          folder_path = "E:/BBK17/pj/data_2023apr/results/bkmr/29_chem/mirec_sex", 
          ln_mixture_vars= chemicals_29,
          groups=chem_group_29,
          outcome_vars= outcomes, 
          covariate_vars= covariates, 
          include_sex = TRUE)

bkmr_cohort(data=dat, 
          folder_path = "E:/BBK17/pj/data_2023apr/results/bkmr/29_chem", 
          ln_mixture_vars= chemicals_29,
          groups=chem_group_29,
          outcome_vars= outcomes, 
          covariate_vars= covariates,  
          include_cohort = FALSE)
```



###############################################################################
###############################################################################

#Extracting PIPs and overall estimates
```{r}
results <- extract_data_from_folder("E:/BBK17/pj/data_2023apr/results/bkmr/29_chem",
                                    select_iterations= seq(25000, 50000, by = 50),
                                    estimate_sequence= seq(0.10, 0.90, by = 0.05))
```

#prepare data for plotting
```{r}
overall_risk<- results[2] |>as.data.frame() |>
  clean_names()|>
  rename("outcome" = "overall_risk_df_measure_name",
         "cohort" = "overall_risk_df_outcome_name",
         "gender"= "overall_risk_df_data_name") |>
  mutate(cohort = fct_recode(as.factor(cohort), Pooled = "all")) |>
  mutate(outcome = fct_recode(as.factor(outcome), FSIQ = "fsiq", VIQ = "viq", PIQ = "piq")) |>
  mutate(gender = fct_recode(as.factor(gender), All = "all", Female = "female", Male = "male"))
```

#plotting overall estimates
```{r}
ggplot(overall_risk, aes(overall_risk_df_quantile, overall_risk_df_est, 
                         ymin = overall_risk_df_est - 1.96*overall_risk_df_sd, 
                         ymax = overall_risk_df_est + 1.96*overall_risk_df_sd)) + 
    geom_hline(yintercept = 00, linetype = "dashed", color = "gray") + 
    geom_pointrange(size = 0.15, aes(color = ifelse((overall_risk_df_est - 1.96 * overall_risk_df_sd) <= 0 &
                                                  (overall_risk_df_est + 1.96 * overall_risk_df_sd) >= 0,
                                                "Credible intervals with null", "Credible intervals without null"))) + 
  scale_color_manual(values = c("Credible intervals with null" = "black", "Credible intervals without null" = "#D55E00")) +
    labs(x=NULL,
       y=expression(paste(beta[italic(bkmr)]," [95% CrI]")),
       title = "",
       caption = "")+
  theme_bw()+
    facet_wrap(gender~outcome, scales = "free")+
    theme(axis.line = element_line(colour = "black"),
          axis.text=element_text(size=10), 
          axis.title=element_text(size=10,face="bold"),
          strip.text = element_text(size=10),
          axis.text.x = element_text(angle = 0, hjust = 1, vjust = 1, size = 10),
          axis.text.y = element_text(size = 10), 
          panel.spacing.x=unit(0.5, "lines"),
          panel.spacing.y=unit(0.5, "lines"),
          panel.grid.minor = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.major.y = element_blank(),
          plot.caption = element_text(size = 6),
          plot.title=element_text(size = 10, hjust = 0),
          panel.border = element_blank(),
          text=element_text(size=10),
          axis.ticks.x=element_blank(),
          strip.text.x=element_text(size=10,angle=0),
          strip.text.y = element_text(size=10),
          strip.placement = "outside",
          strip.background = element_rect(fill =  "transparent"),
          legend.box.just = "center",
          legend.spacing = unit(0, "cm"),
          legend.position = "bottom",
          legend.box = "horizontal",
          legend.direction = "horizontal",
          legend.text = element_text(size = 8),
          legend.key = element_blank(),
          legend.key.height = unit(3, "mm")) +
  guides(col=guide_legend(title = "Statistical significance"))+
  scale_x_continuous(breaks = seq(0.2, 0.8, by = 0.3),
                     labels = seq(0.2, 0.8, by = 0.3))

```

#save effect estimate plots
```{r}
ggsave("E:/BBK17/pj/data_2023apr/results/bkmr/29_chem/29_chem.tiff", 
       width = 10,height = 6,
       dpi=300)
```


# extract pips
```{r}
pip_df<- extract_pips("E:/BBK17/pj/data_2023apr/results/bkmr/9_chem")
write_csv(pip_df, "E:/BBK17/pj/data_2023apr/results/bkmr/9_chem/9chem_pips.csv")
```


#Publishing pips shiny object
# Shiny data format
```{r}
pos_pip_df<- read_csv("E:/BBK17/pj/data_2023apr/results/bkmr/lasso_pos14chem/lasso_14chem.csv") |>
  mutate(chemical_set="LASSO_PCS")

neg_pip_df<- read_csv("E:/BBK17/pj/data_2023apr/results/bkmr/lasso_15chem/lasso_neg_pips.csv")|>
  mutate(chemical_set="LASSO_NCS")

all_chem_pip_df<- read_csv("E:/BBK17/pj/data_2023apr/results/bkmr/29_chem/29_chem.csv") |>
  mutate(chemical_set="All_29_chemicals")

pip_df<- bind_rows(pos_pip_df, neg_pip_df, all_chem_pip_df) |>
  mutate(comps = str_remove(file_name, "bkmr_")) |>
  separate(comps, c("gender", "cohort", "wppsi", "outcome"), sep = "_") |>
  mutate(outcome=toupper(outcome),
         cohort = recode(cohort, "all" = "Pooled","home" = "HOME","mirec" = "MIREC"),
         gender = recode(gender, "all" = "Pooled","male" = "Male","female" = "Female"))|>
  rename(Chemical_group = "group", Environmental_chemical = "variable")

remove(pos_pip_df, neg_pip_df, all_chem_pip_df)

write_csv(pip_df, "E:/BBK17/pj/wppsi_bkmr_pips/bkmr_pips.csv")
```



#for description the paper
```{r}
overall_risk<- results[2] |> as.data.frame() |>
  clean_names() |>
  mutate(overall_risk_df_est = round(overall_risk_df_est, 2),
         ymin = round(overall_risk_df_est - 1.96*overall_risk_df_sd, 2), 
         ymax = round(overall_risk_df_est + 1.96*overall_risk_df_sd, 2)) |>
  select(c(1,2,5,6,8:10))


```


################################################################################
################################################################################
#manual bkmr for PIQ in males
```{r}
dat<- dat |>
  filter(sex=="1")

chemicals_9<- c("Hg", "P_PB",
                 "PBDE_47",  "PFHxS", "PFOA", "PFOS",
                 "BDCIPP", "MBP", "MCPP")

chem_group_9<- c(rep(1,times=2), rep(2,times=4), rep(3, times=3))
outcomes<-c( "wppsi_piq")
covariates<- c("cotinine", "home_score_total", "mom_age", "cohort","city",
               "race_bin", "mom_edu_cat", "parity_n", "fish_int_pm")

# for all chemicals select 1:32
dat_set <- dat |>
  select(c(chemicals_9, covariates, outcomes)) |>
  drop_na()|>
  mutate_at(vars(1:10), ~log10(.+0.0000001)) |>
  drop_na() |>
  mutate_all(as.numeric)

#table(dat_set$sex, dat_set$cohort)
```
#manual bkmr run for PIQ in males
```{r}
bkmr_cohort(data=dat_set, 
          folder_path = "E:/BBK17/pj/data_2023apr/results/bkmr/9_chem", 
          ln_mixture_vars= chemicals_9,
          groups=chem_group_9,
          iter = 50000,
          outcome_vars= outcomes, 
          covariate_vars= covariates,  
          include_cohort = FALSE)
```

#getting additional plots
```{r}
load("E:/BBK17/pj/data_2023apr/results/bkmr/9_chem/bkmr_all_all_wppsi_piq.rda")
```

#univariate plot
```{r}
pred.resp.univar <- PredictorResponseUnivar(fit = bkmr)

ggplot(pred.resp.univar, aes(z, est, ymin = est - 1.96*se, ymax = est + 1.96*se)) + 
    geom_smooth(stat = "identity", color="black", fill="grey85", linewidth=0.5) + 
    facet_wrap(~ variable, scales = "free") +
  ylab("h(z)")+ 
  theme_bw()+
  theme(axis.line = element_line(colour = "black"),
          axis.text=element_text(size=10), 
          axis.title=element_text(size=10,face="bold"),
          strip.text = element_text(size=10),
          axis.text.x = element_text(angle = 0, hjust = 1, vjust = 1, size = 10),
          axis.text.y = element_text(size = 10), 
          panel.spacing.x=unit(0.5, "lines"),
          panel.spacing.y=unit(0.5, "lines"),
          panel.grid.minor = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.major.y = element_blank(),
          plot.caption = element_text(size = 6),
          plot.title=element_text(size = 10, hjust = 0),
          panel.border = element_blank(),
          text=element_text(size=10),
          axis.ticks.x=element_blank(),
          strip.text.x=element_text(size=10,angle=0),
          strip.text.y = element_text(size=10),
          strip.placement = "outside",
          strip.background = element_rect(fill =  "transparent"),
          legend.box.just = "center",
          legend.spacing = unit(0, "cm"),
          legend.position = "bottom",
          legend.box = "horizontal",
          legend.direction = "horizontal",
          legend.text = element_text(size = 8))
```

#interaction plot
```{r}
ln_mixture <- as.matrix(dat_set[, chemicals_9])

pred.resp.bivar <- PredictorResponseBivar(fit = bkmr, min.plot.dist = 1)

pred.resp.bivar.levels <- PredictorResponseBivarLevels(
  pred.resp.df = pred.resp.bivar, Z = ln_mixture, qs = c(0.1, 0.5, 0.9))


ggplot(pred.resp.bivar.levels, aes(z1, est)) + 
    geom_smooth(aes(col = quantile), stat = "identity") + 
    facet_grid(variable2 ~ variable1) +
    ggtitle("h(expos1 | quantiles of expos2)") +
    xlab("expos1")+ 
    theme_bw()+
    theme(axis.line = element_line(colour = "black"),
          axis.text=element_text(size=10), 
          axis.title=element_text(size=10,face="bold"),
          strip.text = element_text(size=10),
          axis.text.x = element_text(angle = 0, hjust = 1, vjust = 1, size = 10),
          axis.text.y = element_text(size = 10), 
          panel.spacing.x=unit(0.5, "lines"),
          panel.spacing.y=unit(0.5, "lines"),
          panel.grid.minor = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.major.y = element_blank(),
          plot.caption = element_text(size = 6),
          plot.title=element_text(size = 10, hjust = 0),
          panel.border = element_blank(),
          text=element_text(size=10),
          axis.ticks.x=element_blank(),
          strip.text.x=element_text(size=10,angle=0),
          strip.text.y = element_text(size=10),
          strip.placement = "outside",
          strip.background = element_rect(fill =  "transparent"),
          legend.box.just = "center",
          legend.spacing = unit(0, "cm"),
          legend.position = "bottom",
          legend.box = "horizontal",
          legend.direction = "horizontal",
          legend.text = element_text(size = 8))
```


################################################################################
################################################################################


