---
title: "mirec_data"
author: "puvvula"
date: "2022-10-05"
output: pdf_document
---

```{r}
library(tidyverse)
library(readxl)
library(janitor)
library(skimr)
library(reshape2)
library(naniar)
```

Outcome: child 3 Yr scores
```{r}
#parent: SRS-2
#child: NEPSY, and WPPSI-3
#get WPPSI composite scores from here
################################################################################
################################################################################
ndev<-  read_csv("E:/BBK17/1. BBK17_June10 2022/CRF/2. MIREC-CD Plus STUDY/Data/NDEV.csv")

ndev <- select(ndev, SUBJECT_ID, AgeTestingMonths, WPPSI1_2, WPPSI2_2, WPPSI3_2, 
               WPPSI4_2, WPPSI5_2, WPPSI6_1, WPPSI7_1, WPPSI8_1, WPPSI9_1)

colnames(ndev) <- c("subject.id", "age", "wppsi.1","wppsi.2", "wppsi.3",
                    "wppsi.4", "wppsi.5", "viq", "piq", "fsiq", "general.language")

#filter composite scores
wppsi<- ndev[c(1,8:10)] |> clean_names()
write_csv(wppsi, "E://BBK17//pj//outcome//wppsi.csv")

################################################################################
################################################################################
#filter basc ext, int & bsi here
dt_out<- read_excel("E://BBK17//1. BBK17_June10 2022//DV//2. MIREC-CD Plus STUDY//BBK17_CDPlus_BABY_DVs.xlsx") |> clean_names() 

basc_brief<- dt_out[c(1,28,29,30,40,42)] |>
  rename(basc_ext=bastexcddv,
         basc_int=bastincddv,
         basc_bsi=bastbscddv,
         brief_wm=brietwmcddv,
         brief_po=brietpocddv) #BSI=behavioral symptom index
write_csv(basc, "E://BBK17//pj//outcome//basc_brief.csv")

################################################################################
################################################################################
# join basc and wppsi
mirec_out<- full_join(basc_brief, wppsi, by="subject_id")
write_csv(mirec_out, "E://BBK17//pj//mirec_out.csv")
```


First trimester - Exposure
```{r}

# triclosan (removing not reported records)
dat<- read_csv2("E://BBK17//3. BBK17_lab test correction on Dacima_Nov7 2022//MIREC_RESEARCHERS RESULTS table_code 001//QueryData_code001.csv") |>
  clean_names() |>
  rename(tests_results_raw_result =researcher_results_raw_result) |>
  filter(complete.cases(researcher_results_unit_set))

dat$tests_results_test_name<- str_extract(dat$researcher_results_test_name, "^.{3}") 
dat$tests_results_test_name<-as.factor(dat$tests_results_test_name)
dat$tests_results_raw_result<-as.numeric(dat$tests_results_raw_result)
dat<- dat |> filter(as.factor(tests_results_test_name) == "TRI")

# sum of triclosan glusuronide, free and sulfate
labres <- dat |> 
  group_by(specimen_bar_code) |>
  summarise(test.result=sum(tests_results_raw_result)) |>
  mutate(tests_results_test_name ="triclosan", visit = "12w")

#flag LT lod values for total triclosan 
# LOD = 0.12; from https://ehp.niehs.nih.gov/doi/epdf/10.1289/ehp.1408187
labres$lod.flag<-  as.numeric(ifelse(labres$test.result< 0.1200, 1, 0))

#Phthalates
labres<- read_csv2("E://BBK17/2. BBK17_pending lab tests_July19 2022//LAB//MIREC_MICROALIQOUT RESULTS table code 001-01.001-03//QueryData_code 001-01.csv") |>
  clean_names()|>
  rename(tests_results_test_name = micro_aliquot_tests_results_test_name,
         tests_results_raw_result =micro_aliquot_tests_results_raw_result)

#BPA, Urine Arsenic, OP pesticides, OPFR, Specific gravity
labres <-read_csv2("E://BBK17//1. BBK17_June10 2022//LAB//MIREC_TEST RESULTS table_codes 001.002.009.013.016.049.054.060.096.097.101//QueryData_code001.csv") |>
  clean_names() |> 
  mutate(id_join = str_sub(specimen_bar_code, 1,6))

df_1a <- df_1[c(5,6,10)] |> 
  pivot_wider(names_from = c(tests_results_test_name),
              values_from = tests_results_raw_result) 

num<- names(df_1a[c(2:15)])
df_1a[num]<- sapply(df_1a[num], as.numeric)

write_csv(df_1a, "E://BBK17//pj//pj_dat//ex_dat6.csv")

#Urine: Phthalates
df_2 <-read_csv2("E://BBK17//1. BBK17_June10 2022//LAB//MIREC_TEST RESULTS table_codes 001.002.009.013.016.049.054.060.096.097.101//QueryData_code002.csv") 

#Urine: Creatinine
df_6 <-read_csv2("E://BBK17//1. BBK17_June10 2022//LAB//MIREC_TEST RESULTS table_codes 001.002.009.013.016.049.054.060.096.097.101//QueryData_code049.csv") |> clean_names()

#Plasma: PBDE, PCB, OC Pesticides
df_3 <-read_csv2("E://BBK17//1. BBK17_June10 2022//LAB//MIREC_TEST RESULTS table_codes 001.002.009.013.016.049.054.060.096.097.101//QueryData_code009.csv") |> clean_names()

#Plasma: Tobacco and PFCs
df_4 <-read_csv2("E://BBK17//1. BBK17_June10 2022//LAB//MIREC_TEST RESULTS table_codes 001.002.009.013.016.049.054.060.096.097.101//QueryData_code013.csv") |> clean_names()

#Whole Blood: - metals
df_5 <-read_csv2("E://BBK17//1. BBK17_June10 2022//LAB//MIREC_TEST RESULTS table_codes 001.002.009.013.016.049.054.060.096.097.101//QueryData_code016.csv") |> clean_names()


#OPFR
labres<- read_csv2("E://BBK17/2. BBK17_pending lab tests_July19 2022//LAB//MIREC_MICROALIQOUT RESULTS table code 001-01.001-03//QueryData_code001-03.csv") |>
    clean_names()|>
    rename(tests_results_test_name = micro_aliquot_tests_results_test_name,
           tests_results_raw_result =micro_aliquot_tests_results_raw_result)

```

Standardizing units
```{r}
#Metals: Lead & Mercury (mol. wt from pubchem)

#molecular weight x conversion factor to get values in mcg/L
pb<- 207.2/1e6 # micromole/L -> mcg/L
hg<-200.59/1e9 #nanomole/L -> mcg/L
#multiply above values with absolute result from imputed dataset
dt<- read_csv("E://BBK17/pj//imputed_data//HGSG.csv")
dt$result.res<-dt$result.res*hg*1e6

write_csv(dt, "E://BBK17/pj//imputed_unit_corr//HGSG.csv")

# PCBs,PBDE, DDE (mcg/L -> ng/L)
dt<- read_csv("E://BBK17/pj//imputed_data//BDE47PL.csv")
dt$result.res <- dt$result.res*1000
write_csv(dt, "E://BBK17/pj//imputed_unit_corr//BDE47PL.csv")

# DMAA micromol/L -> PPB (mol. wt from pubchem)
dt<- read_csv("E://BBK17/pj//imputed_data//DMAAUR.csv")
dt$result.res1 <- (dt$result.res*170)
write_csv(dt, "E://BBK17/pj//imputed_unit_corr//DMA.csv")

```

specific gravity standardization = ((analyte concentration x (median urine specific grav. -1)) / (1-individual specific grav. -1))
Creatinine adjustment = analyte concentration / creatinine concentration
lipid adjustment = (analyte concentration / lipid level)*102.6

```{r}
urine_sg<- read_csv("E://BBK17/pj//imputed_unit_corr//DENS.csv")
urine_sg$join_id<- str_extract(urine_sg$subject_id,  "^.{8}")

dt<- read_csv("E://BBK17/pj//imputed_unit_corr//triclosan.csv") 
dt$join_id<- str_extract(dt$subject_id,  "^.{8}")
dat<- inner_join(dt,urine_sg, by="join_id")

med_sg<- median(dat$result.res.y)
dat$tcs_sg <- ((dat$result.res.x*(med_sg-1))/(dat$result.res.y-1))

dat<- dat[c(1,2,10)] |> rename(tcs = result.res.x, subject_id = subject_id.x) |>
  pivot_longer(!subject_id, names_to = "analyte", values_to = "conc")

write_csv(dat, "E://BBK17/pj//imputed_unit_corr//sg_adj//tcs.csv")


################ creatinine adjustment ########################

cre<- read_csv("E://BBK17/pj//imputed_data//CRE.csv")
cre$join_id<- str_extract(cre$subject_id,  "^.{8}")

dt<- read_csv("E://BBK17/pj//imputed_data//COTISE.csv") 
dt$join_id<- str_extract(dt$subject_id,  "^.{8}")
dat<- inner_join(dt,cre, by="join_id")

dat$cotinine_cr <- (dat$result.res.x/dat$result.res.y)
dat<- dat[c(1,2,10)] |> rename(cotinine = result.res.x) |>
  pivot_longer(!subject_id.x, names_to = "analyte", values_to = "conc")

write_csv(dat, "E://BBK17/pj//imputed_data//cre_adj//cotinine.csv")

############ lipid adjustment for DDE, PBDE, PCB ############

lipi<-  read_csv("E://BBK17/pj//imputed_unit_corr//LIPIPL.csv") 
lipi$join_id<- str_extract(lipi$subject_id,  "^.{8}") 

dt <- read_csv("E://BBK17/pj//imputed_unit_corr//BDE47PL.csv") 
dt$join_id<- str_extract(dt$subject_id,  "^.{8}")
dat<- inner_join(dt,lipi, by="join_id")
 
dat$pbde47_lp <- (dat$result.res.x/dat$result.res.y)
dat<- dat[c(1,2,10)] |> rename(pbde47 = result.res.x, subject_id = subject_id.x) |>
  pivot_longer(!subject_id, names_to = "analyte", values_to = "conc")

write_csv(dat, "E://BBK17/pj//imputed_unit_corr//lipi_adj//pbde47.csv")

```

Converting all exposures to long format - join
```{r}
dat<- read_csv("E://BBK17/pj//imputed_data//HGSG.csv") 
dat<- dat[c(1,2)] |> rename(hg = result.res) |>
  pivot_longer(!subject_id, names_to = "analyte", values_to = "conc")

write_csv(dat, "E://BBK17/pj//imputed_data//merging//hg.csv")
```

Merging exposures to long format
```{r}
file_names<- list.files(path= "E://BBK17//pj//imputed_unit_corr//mirec_fin_merge",
                        recursive = T,
                        pattern = "\\.csv$",
                        full.names = T)

dat<- read_csv(file_names)
dat$subject_id<- str_extract(dat$subject_id,  "^.{6}")

write_csv(dat, "E://BBK17//pj//mirec_exposure_fin.csv")

######## Exposure data summary ####################
geomean <- function(x, na.rm = T, trim = 0, ...) 
  { exp(mean(log(x, ...), na.rm = na.rm, trim = trim, ...)) }

geosd <- function(x, na.rm = T, ...)
  { exp(sd(log(x, ...), na.rm = na.rm, ...)) }

dat_summ<- dat |> 
  group_by(analyte) |>
  summarise(no_obs= n(),
            geometric_mean = geomean(conc), geometric_SD = geosd(conc),
            arithmatic_mean=mean((conc), na.rm=T), arithmatic_std=sd((conc), na.rm = T),
            median = median((conc), na.rm = T), q25 = quantile(conc, .25, na.rm=T),
            q75 = quantile(conc, .75, na.rm=T)) |>
  mutate_if(is.numeric, round, 3)

write_csv(dat_summ, "E://BBK17//pj//mirec_exposure_summary.csv")

```


Exposure missing data summary
```{r}
dat<- read_csv("E://BBK17//pj//mirec_exposure_fin.csv") |>
  pivot_wider(names_from = analyte, values_from = conc)

# filter sg/lp adjusted variables
dat<- dat[c(1,3,5,7,9,11,13,15,17,19,21,23,25,26,28,30,32,34,36,38,40,42,44,46,47,49,51,53,55:58,60)] 
dat$site<- str_extract(dat$subject_id,  "^.{2}") |>
  as.factor()

dat<- dat |> drop_na()

#Select either WPPSI or BASC from here (dont use both outcomes from here)
dat_out<- read_csv("E://BBK17//pj//mirec_out.csv")

gg_miss_upset(dat_out, nsets=5, nintersects=NA)

dat_out$site<- str_extract(dat_out$subject_id,  "^.{2}") |>
  as.factor()

#site specific count by outcomes

x<-dat_out[c(1:6,10)] |>
  drop_na()

```

combine exposure and outcome
```{r}
dat_fin<- inner_join(dat, dat_out, by="subject_id") |>
  drop_na()

cormat<- round(x=cor(dat_fin[2:40], method = "spearman", 
                     use= "complete.obs"), digits=2) |>
  melt() |> clean_names()

ggplot(cormat, aes(x=var2, y=var1, fill=value)) +
  geom_tile(color="white")+
  scale_fill_gradient2(low="red", high="blue", mid="white",
                       midpoint=0,
                       limit=c(-1,1), space= "Lab",
                       name="Spearman\nCorrelation")+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 12, hjust = 1),
        axis.text.y = element_text(angle = 0, vjust = 0.15, size = 12, hjust = 1),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        axis.ticks = element_blank())+
  coord_fixed()

ggsave("E://BBK17//pj//corrplot.tiff",
       width=14, height= 12, dpi=300)
```



